### 2014-12-13 paper.tex
Add more explanatory text to moment metrics

  The very simplest metrics for compactness simply measure the length and width of the district as compared to that of a circle or rectangle of the same area.  There are several variations on this method, but they are all fairly crude.  Other metrics compare the area of the district to the area of the smallest circumscribing circle, or the largest inscribed circle, or to the convex hull of the district -- the smallest convex shape containing the entire district.  Niemi~et~al.~\cite{niemi} list a number of variations on these metrics.  An alternate method of computing dispersion is by computing the moment of inertia of the area of the district.  The concept is borrowed from physics, and is defined as
  \[I = \iint \left((x-x_0)^2 + (y-y_0)^2\right) dx\, dy \]
  where $(x_0, y_0)$ is the center of mass of the district (weighted by area).  More dispersed districts have higher moments of inertia, because the average square distance of a point in the district to the center is larger.  In order to ensure that the metric does not depend on the size of the state, and to make it range from 0 to 1 with 0 being the least compact, we in fact use the measure $A/\sqrt{2\pi I}$.  This metric seems to have been first proposed by Kaiser.~\cite{kaiser}

@@

  Lastly, since ultimately it is not the geography that matters, but the people that live in it, we can modify some of these metrics to take into account population.  Those that compare the area of the district to that of the smallest circumscribing circle or that of the convex hull of the district may simply compare population instead of area, as proposed by Hofeller and Grofman.~\cite{hofeller}.  Alternately, instead of computing the area-weighted moment of inertia, one can compute the population-weighted moment of inertia, which is defined as
  \[I_P = \iint \left((x-x_0)^2 + (y-y_0)^2\right) \rho\, dx\, dy \]
  where $\rho$ is the population density and $(x_0, y_0$ is the center of mass of the district, now weighted by population.  In practice, since we do not know the location of each individual in the district, we may simply assume that each person lives at the center of the census tabulation block in which they were counted.  Then the formula becomes
  \[I_P = \sum_{b} \left((x_b-x_0)^2 + (y_b-y_0)^2\right) P_b\]
  where the sum is taken over census tabulation blocks $b$, with $(x_b, y_b)$ the center of mass (weighted by area) of the census block, and $P_b$ its population.  In order to make the metric less dependent on population, we can use the metric $AP_d/(2\pi I_P)$; this does not necessarily normalize the metric to be between 0 and 1, but at least makes it independent of the total area and total population.  Using moment of population to measure district compactness was first proposed by Weaver.~\cite{weaver}

>>>>

  The very simplest metrics for compactness simply measure the length and width of the district as compared to that of a circle or rectangle of the same area.  There are several variations on this method, but they are all fairly crude.  Other metrics compare the area of the district to the area of the smallest circumscribing circle, or the largest inscribed circle, or to the convex hull of the district -- the smallest convex shape containing the entire district.  Niemi~et~al.~\cite{niemi} list a number of variations on these metrics.  An alternate method of computing dispersion is by computing the moment of inertia of the area of the district.  The concept is borrowed from physics, in which rotating bodies have a moment of inertia that defines how their mass is distributed: a bicycle wheel has a higher moment of inertia than a ball of the same mass because its mass is further from its axis, and thus the wheel is more inclined to keep rotating once started.  The moment of area simply assumes that the district has equal ``mass'' everywhere, and is defined as
  \[I = \iint \left((x-x_0)^2 + (y-y_0)^2\right) dx\, dy \]
  where $(x_0, y_0)$ is the center of mass of the district (weighted by area).  More dispersed districts have higher moments of inertia, because the average square distance of a point in the district to the center is larger.  In order to ensure that the metric does not depend on the size of the state, and to make it range from 0 to 1 with 0 being the least compact, we in fact use the measure $A/\sqrt{2\pi I}$.  This metric seems to have been first proposed by Kaiser.~\cite{kaiser}

@@

  Lastly, since ultimately it is not the geography that matters, but the people that live in it, we can modify some of these metrics to take into account population.  Those that compare the area of the district to that of the smallest circumscribing circle or that of the convex hull of the district may simply compare population instead of area, as proposed by Hofeller and Grofman.~\cite{hofeller}.  Alternately, instead of computing the area-weighted moment of inertia, one can compute the population-weighted moment of inertia, using the population as the analogy of ``mass''.  Formally, the moment of population is defined as
  \[I_P = \iint \left((x-x_0)^2 + (y-y_0)^2\right) \rho\, dx\, dy \]
  where $\rho$ is the population density and $(x_0, y_0$ is the center of mass of the district, now weighted by population.  In practice, since we do not know the location of each individual in the district, we may simply assume that each person lives at the center of the census tabulation block in which they were counted.  Then the formula becomes
  \[I_P = \sum_{b} \left((x_b-x_0)^2 + (y_b-y_0)^2\right) P_b\]
  where the sum is taken over census tabulation blocks $b$, with $(x_b, y_b)$ the center of mass (weighted by area) of the census block, and $P_b$ its population.  In order to make the metric less dependent on population, we can use the metric $AP_d/(2\pi I_P)$; this does not necessarily normalize the metric to be between 0 and 1, but at least makes it independent of the total area and total population.  Using moment of population to measure district compactness was first proposed by Weaver.~\cite{weaver}
